### 2019-11-19 tex/Tracking.tex
proof of concept and application section updated

\documentclass[3p,times]{elsarticle}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{algorithm2e}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}

\usepackage{xcolor}
\usepackage[version=4]{mhchem}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\diag}{\textbf{diag}}
\newtheorem{defn}{Definition}
\newtheorem{hypo}{H}
\newtheorem{lemma}{Lemma}
\newtheorem{theo}{Theorem}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

\graphicspath{{Images/}}
% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

\title{A general model of microbial interactions in a chemostat.}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author{Pablo Ugalde-Salas, Jérôme Harmand, Elie Desmond Le-Quemener, Héctor Ramirez}

\address{}

\begin{abstract}
	A model of microbial interactions applied to a nitrifying process in a chemostat setting is revisited. To overcome the shortcomings of this earlier model an optimal tracking method is developed to include the genetic measurements for the prediction of a bioreactor. The resulting model is a high order (24 variables) differential equations system with a good prediction of measured metabolites. Numerical simulations of the model are presented and discussed.
\end{abstract}

\end{frontmatter}

\section{Introduction}

Microbial communities and their interactions play a central role in the understanding of microbial ecosystems \cite{West2007}, and a current challenge is integrating genetic sequencing data in a deterministic modelling framework \cite{Widder2016,WADE201664}. A thorough review in current methodologies in the deterministic modelling approaches of microbial community dynamics is presented by Song \textit{et al.} \cite{Song2014}. Using their terminology this articles deals with population-based approaches where species are taken as the interacting units.

The classical ecological concept of species and niche in the microbial world is an elusive one: In the macro world one can clearly identify one species of another for reproductive reasons and their ability to give birth to offspring. In the case of bacteria and archea, reproduction goes simply by mitosis and exchange of some functional genes (the ability to synthesize or metabolize substances) can be acquired in evolutionary scale through lateral gene transfer \cite{Boucher2003}. Therefore as an ecological problem is hard to precise the 'niche' of 'species'.

These obstacles can be circumvented by considering the microbiologist concept of operative taxonomic unit (OTU) based in 16sRNA analysis. There is a core genome in Bacteria, which is in turn related to the main metabolism of Bacteria. The abundance of the present OTU allows to account for a group of Bacteria that behave similarly in terms of which substances catabolyzes and at which growth rates and substrate to biomass conversion yields \cite{Muyzer1993}.

In the past years considerable efforts have been made to measure the bacterial community composition. Tests such as fluorescence in situ hybridization, polymerase chain reaction (PCR) dependent techniques, and PCR independent techniques for the analysis of DNA, RNA, proteins, and metabolites have become a standard tool for studying  microbial diversity\cite{FERRERA2016790}.

The main contribution of this article is a procedure to integrate the microbial community measurements in chemostat models, based on any technique that can quantify the species abundances through time. This is a first step towards developing predictive models of a reactor's performance based on the microbial community structure.

As a departing point the work of Dumont et al.\cite{Dumont2016} is reinspected in section 2. They model a chemostat experiment where nitrification takes place by considering a generalized Lotka-Volterra (gLV) model \cite{Hernandez-Bermejo1997} coupled with a substrate limited growth expression. Section 3 deals with the mathematical analysis of their model. It is shown that the number of possible equilibrium points grows exponentially with the number of species. Finally after a numerical analysis it is conjectured that there always exists a single stable equilibrium point depending on the operational parameters.

Section 4 is the article's most original contribution. The concept of interaction function is developed such that it generalizes the gLV model. For the identification of the interaction function a method of optimal tracking was adapted: The growth rate of each species is modulated by a constrained regular control of the system. The regular control is composed of a feedback part on the species state variable, and a feed forward part, or tracking, on the measurement of abundances of each species.

In section 5 the methodology from section 4 is applied to the data from the experiments perform by Dumont et al \cite{Dumont2009}. Finally, a discussion on the scope of the method, its shortcomings and perspectives is presented.

\section{Model Definition}

Notations used throughout the article are:
\begin{enumerate}
	\item $n$: the number of $OTU$ considered.
	\item $n_i, \, i \in \{1,2\}$: the number of OTU in functional group $G_i$.
	\item Let $m$ be an interger then $[m]:=\{1,\dots,n \}$. 
	\item $x_i$ :  is the concentration of OTU $i$ measured in $[g/l]$. $i \in [n]$. 
	\item $x$: vector $(x_1,\dots,x_n)^\top$.
	\item $s_1$ : concentration of substrate 1 in $[g/l]$.
	\item $s_2$ : concentration of substrate 2 in $[g/l]$. 
	\item $s_3$ : concentration of substrate 3 in $[g/l]$.
	\item $s$:  vector $(s_1,s_2,s_3)$.
	\item $\I_i(t,x)$ : Interaction function of OTU $i \in \{1,\dots,n\}$.
	\item $\mu_i(s,x)$ : growth function of OTU $i \in \{1,\dots,n\}$.
	\item $\mu = (\mu_1(s,x), \dots, \mu_n(s,x))$ vector containing the growth function of every OTU.
	\item $D$: Dilution Rate of the continuous reactor in $[1/day]$. 
	\item $y_i$: Yield of moles of biomass formed per mole of substrate consumed of OTU $i$.
	\item $y$: vector $y_1,\dots,y_n$.
	\item $k_i$: Yield of  moles of substrate consumed per mole of biomass formed of OTU $i$ . The inverse of the biomass yield $k_i = \dfrac{1}{y_i}$.
	\item $k$: vector $k_1,\dots,k_n$.
	\item For integers $m_1$ and $m_2$ and $a \in \R$, $a_{m_1 \times m_2}$ represents a matrix of $m_1$ rows and $m_2$ columns with $a$ in every entry. 
	\item Let $m$ be an integer then $I_m$ is the identity matrix of size $m$.
\end{enumerate}

\subsection{Stoichiometric Equations}
A cascade (bio)reaction process is considered. Suppose $n$ different OTU are present in the chemostat. The cascade reaction refers to the situation where a group of microorganisms ($G_1 \subset [n] $) consumes a substrate $s_1$ and produces $s_2$ and biomass, while another group of microorganisms ($G_2\subset [n]$) consumes $s_2$ and produces $s_3$ and biomass. $G_1$ and $G_2$ are called functional groups. The number of organisms in each functional will be denoted $n_1$ and $n_2$ respectively, that is $\vert G_1 \vert = n_1$ and $\vert G_2 \vert = n_2$ . This work treats the case when the $G_1$ and $G_2$ are disjoint sets:

\begin{hypo}  Sets $G_1$ and $G_2$ satisfy: $G_1 \cap G_2 = \emptyset$ and $G_1 \cup G_2 = [n]$. 
\end{hypo}

The situation is described as simplified reactions \eqref{Reaction1} and \eqref{Reaction2}. The reactions are simplified in the sense that they do not attempt to represent a balanced chemical reaction, rather it represents the direction of the bioprocess and the proportions of different consumed and formed compounds of interest.

The terms $y_i$ are known as yields, it represents the number of grams of biomass produced per gram of substrate consumed. In this work the unit grams of dry biomass per gram of substrate is used for yields. Furthermore, for each $ i \in [n]$, OTU $i$ is characterized by its process rate (also known as growth function) $\mu_i(s,x)$. Notice that for being as generic as possible, the growth rate may be a function of the whole state in order to model the influence of all OTU on the growth rates of others.

\begin{align}
\label{Reaction1} \tag{R G1} 	s_1 \stackrel{\mu_i(s,x)}{\longrightarrow} s_2 + y_ix_1 \quad \forall i \in G_1 \\
\label{Reaction2} \tag{R G2} s_2\stackrel{\mu_i(s,x)}{\longrightarrow} s_3 + y_ix_i \quad \forall i \in G_2
\end{align}

\subsection{Kinetic Equations}
The growth rate of each species is given by:
\begin{align}
\mu_i(s,x) &= \bar{\mu}_i \dfrac{s_1}{K_i + s_1}\left(1+\sum \limits_{j\in [n]} a_{ij} x_{j} \right) \quad \forall i \in G_1 \\
\mu_i(s,x) &= \bar{\mu}_i \dfrac{s_2}{K_i + s_2}\left(1+\sum \limits_{j\in [n]} a_{ij} x_{j} \right) \quad \forall i \in G_2
\end{align}
The term $\left(1+\sum \limits_{j\in [n]} a_{ij} x_{j} \right)$ accounts for pairwise interactions affecting the growth rate of each OTU. Let $A$ denote the matrix
with entries $a_{ij}$, which will be referred to as the interaction matrix.

\subsection{Mass Balance Equations}

Consider the scenario of an continuous and homogeneous reactor: the input flow is the same as the output flow, with a dilution rate of $D$. The input flow contains a concentration $s_{in}$ of substrate $s_1$. Note that $s_{in}$ and $D$ are time dependent. Each OTU grows at a rate $\mu_i(s,x)$. System \eqref{system} represents this situation. A specific case of $\mu_i(s,x)$ is given by \cite{Dumont2016}, and the first part of this article deals with the mathematical analysis of that model. A second part then generalizes its growth function and calibrates it against experimental data using an optimal control approach.

\begin{align} 
\label{system}
\begin{array}{cl}
\dot{x_i} =& \left(\mu_i(s,x) -D \right)x_i \quad \forall i \in G_1\\
\dot{x_i} =& \left(\mu_i(s,x) -D \right)x_i \quad \forall i \in G_2 \\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D-\sum\limits_{i\in G_1} \frac{1}{y_i}\mu_i(s_1,x) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i\in G_1}\frac{1}{y_i}\mu_i(s,x)x_i 		-\sum\limits_{i\in G_2}\frac{1}{y_i}\mu_i(s,x) x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i\in G_2}\frac{1}{y_i}\mu_i(s,x) x_i 
\end{array}
\end{align}

\section{Mathematical Analysis}

The system of equations \eqref{system} is defined in the region

\begin{align*}
\Omega := \{ (x_1,\dots,x_n,s_1,s_2,s_3)\in  \R^{n+3} | x_1,\dots,x_n,s_1,s_2,s_3  \geq 0  \}
\end{align*}

First sufficient conditions for the system to be well posed are established: meaning that solutions remain bounded and non-negative in time, this ultimately implies that the solution exists for every $t \geq 0$ \cite{Khalil1996}.

Second the equilibria of the system are derived. Possible equilibrium points for this system grow exponentially with the number of OTU considered ($n$).  Stability is not analytically addressed, a numerical scheme calculating every equilibrium point and the system's Jacobian eigenvalues at the equilibrium point were implemented for studying the system.

\subsection{Properties of the system}

 By bounding the norm of the interaction matrix on the initial conditions and parameters one establishes that solutions will remain positive and bounded.

\begin{lemma}
	\label{l1}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists positive scalars $M_1$, $M_2$, and $M_3$ such that solutions to \eqref{system} satisfy the following inequalities:
	\begin{align}
	&\sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1 \leq M_1 \\
	&\sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 +s_2 \leq M_2 \\
	&s_1 + s_2 + s_2 \leq M_3
	\end{align}
\end{lemma}

The proof can be seen in the appendix. A bound on the norm of $A$ is found such that every matrix $A$, respecting the bound, guaranties that $\Omega$ is a positively invariant set.

\begin{theo}
	\label{theoWellPosedness}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists a constant $M>0$ such that for every matrix $A$ satisfying $\Vert A \Vert_{\infty} \leq M $, the solutions remain in $\Omega$ and are bounded.
\end{theo}

The proof can be seen in the appendix.

\subsection{Equilibrium Points}

In this section analytical expressions for equilibrium points are derived. Analytic expressions of equilibrium points are obtained, however no analytic expression concerning the stability of such points is presented. To answer the question of stability a numeric scheme is used by evaluating the Jacobian at the equilibrium point. At the end of the section an algorithm is provided for exploring all the possible equilibria.

\begin{defn} 
	An equilibrium point (or steady state) is a point $(x^{eq},s^{eq}) \in \R^{n+3}_+$ so that the right hand side of equations \eqref{Eq1},\eqref{Eq2}, \eqref{Eq3}, and \eqref{Eq4} equals zero. 
\end{defn}

Observe that equilibrium points are by definition non-negative so the state variables can have biological meaning. Details of the calculations are given in the appendix.

This model \eqref{system} is restated in a more compact form using matrix notation.

\begin{defn} Given a vector $v=(v_1,\dots, v_n)\in \R^n$, the function $\diag(v)$ stands for:
	\begin{align}
	\begin{array}{rc}
	\diag:\R^n \rightarrow & \mathbb{M}_{n\times n}(\R)\\
	& \\
	v \rightarrow & \begin{pmatrix}
	v_1 & 0 & \dots & 0 \\
	0 & v_2 & \ddots & \vdots\\ 
	\vdots & \ddots & \ddots& 0 \\
	0 & \dots &0 & v_n 
	\end{pmatrix}
	\end{array}
	\end{align}
\end{defn}

The former function places a vector $v$ in the diagonal of a matrix. By noting $A$ the matrix with coefficients $a_{ij}$, system \eqref{system} is restated as follows:
\begin{align}
\label{Eq1}	\dot{x} = & \diag(x)(\diag(f(s))(1_{n\times 1} + Ax) - D_{n\times 1}) \\
\label{Eq2}	\dot{s_1} = & (s_{in}-s_1)D + \begin{bmatrix}
-(1_{n_1\times 1} +A_1x) \\0
\end{bmatrix}^\top \diag(k) \diag(f(s))x \\
\label{Eq3}	\dot{s_2} = & -s_2D+
\begin{bmatrix}
1_{n_1\times 1} +A_1x \\ -(1_{n_2\times 1} +A_2x)
\end{bmatrix}^\top \diag(k) \diag(f(s))x \\	
\label{Eq4}	\dot{s_3} = & -s_3D+\begin{bmatrix}
0	\\(1_{n_2\times 1} +A_2x) 
\end{bmatrix}^\top \diag(k) \diag(f(s))x 
\end{align}

%
%The following notation was used:
%$A = \begin{bmatrix}
%A_1\\A_2
%\end{bmatrix}$.
%
%Where $A_1$ is a matrix of dimension $n_1 \times n$ corresponding to the interaction coefficients of OTU $i \in G_1$. $A_2$ is a matrix of dimension $n_2 \times n$ rows corresponding to the interaction coefficients of OTU $i \in G_2$.

\subsubsection{Complete Coexistence}

This section derives the equations for an equilibrium point with every OTU being positive: $x^{eq}_i>0$.

From equation \eqref{Eq1} one gets that an equilibrium point satisfies: \begin{align*}
\diag(x)(\diag(f(s)(1_{n\times 1} + Ax) - D_{n\times 1}) & = 0 \\
\Rightarrow x_i = 0 \vee (\diag (f(s))(1_{n\times 1} + Ax) - D_{n\times 1})_i & = 0 \quad \forall i\in (1,\dots,n)
\end{align*}

Since $x_i>0$ one has 
\begin{align}
\diag(f(s))(1_{n\times 1} + Ax) - D_{n\times 1}=0 \\
\label{xEquilibria}\Leftrightarrow x = A^{-1}(\diag (f(s))^{-1}D_{n\times 1} - 1_{n\times 1})
\end{align}

By replacing \eqref{xEquilibria} in equation \eqref{Eq2}:

\begin{align}
\label{s1Equilibria} (s_{in}-s_1)-	\begin{bmatrix}
1_{n_1\times 1} \\0_{n_2\times 1}
\end{bmatrix}^\top \diag (k) A^{-1}(\diag(f(s))^{-1}D_{n\times 1}-1_{n\times 1}) = 0 
\end{align}

A similar expression is obtained when replacing \eqref{xEquilibria} in equation \eqref{Eq3}

\begin{align}
\label{s2Equilibria}-s_2+
1_{n\times 1}^{\top}\diag (k)  A^{-1}(\diag(f(s))^{-1}D_{n\times 1}-1_{n\times 1}) = 0
\end{align}

From equation \eqref{s1Equilibria} $s_2$ can be written as a function of $s_1$:
\begin{align}
s_2 = \frac{s_1}{c_1s_1^2+c_2s_1+c_3} \label{S2(S1)}
\end{align}

Then by replacing  \eqref{S2(S1)} in equation \eqref{s2Equilibria}, one gets a fourth degree polynomial for $s_1$.

\begin{align}
\label{Poly4} a_4s_1^4+a_3s_1^3+a_2s_1^2+a_1s_1+a_0 = 0
\end{align}

Formulae for coefficients $c_1, c_2, c_3, a_0, a_1, a_2, a_3, a_4$ can be found in the appendix.

Finally, in equilibrium, $s_3$ can be defined in terms of $s_1$, $s_2$ and $s_{in}$. This is done by adding equations \eqref{Eq2}, \eqref{Eq3},  and \eqref{Eq4} which gives :

\begin{align}
s_{in} = s_1+s_2+s_3 \label{s1s2s3}
\end{align}

If the system of equations \eqref{xEquilibria}, \eqref{S2(S1)}, \eqref{Poly4}, and \eqref{s1s2s3} provides only solutions with some negative entries then complete coexistence is not possible.

\subsubsection{Washout of some OTU}

\begin{defn}
	Given an equilibria point $(x^{eq},s^{eq})$ of system \eqref{system}, then the set of non-active coordinates $\mathcal{J}\subset \{1,\dots,n\}$ is defined as:
	$\mathcal{J} = \{j_1, \dots, j_m\ : \, x^{eq}_{j_i} = 0 , \;i \in [m] \} $.

	The active point $x^{act}\in \R^{n-m}$ is defined by the positive entries of $x^{eq}$. The relationship between the two is given by:

	\begin{align}\label{EquilibriaFormula} (x^{eq})_i = \begin{cases}
	x^{act}_i & \text{for } i < j_1\\
	x^{act}_{i-k} & \text{for } j_{k}< i <j_{k+1}  \text{ and } k \in \{1,\dots,m-1\}\\
	x^{act}_{i-m} & \text{for } j_{m}< i \\
	0 & \text{for } i \in \{j_1, \dots, j_m\}
	\end{cases} \end{align}

	The active growths $f^{act}(s)$ and yields $k^{act}$ are defined analogously by the positive entries of $x^{eq}$. The active interactions $A^{act}$ is defined as matrix $A$ without the $\mathcal{J}$ rows and columns.
\end{defn}

From equation \eqref{Eq1} equilibrium points satisfy:

\begin{align}
\Rightarrow  (\diag (f(s))(1_{n\times 1}+ Ax) - D_{n\times 1})_i = 0 \quad \forall i \in [n]\setminus \mathcal{J} \\
\Leftrightarrow f_i(s)(1 + A_{i\bullet}x) - D = 0 \quad \forall i \in [n]\setminus \mathcal{J} \label{Extinction}
\end{align}

But since $\forall j \in \mathcal{J}, \; x^{eq}_j = 0 $ then the coefficients $a_{ij}$ play no role in the equation \eqref{Extinction} one can rewrite equation \eqref{Extinction} as $\diag(f^{act}(s))(1_{(n-m)\times 1} + A^{act}x^{act}) - D_{(n-m) \times 1}=0 $.

\begin{lemma} 
	The active point of the solution of \eqref{Extinction} solves the system of equations \eqref{EqSome}. 
	\begin{align}
	\label{EqSome}x^{act} = (A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1})
	\end{align} 
\end{lemma}

The following hypothesis represents the case where in each functional group remains at least one OTU :
\begin{hypo}The set $\mathcal{J}$ satisfies $ G_1 \not \subset \mathcal{J} , \, G_2 \not \subset \mathcal{J} $.
	\label{noWashoutHyp}
\end{hypo}

Under H \ref{noWashoutHyp}, the equilibrium of the system is solved as described by equations \eqref{S2(S1)}, \eqref{Poly4}, and \eqref{s1s2s3} as before, but coefficients are calculated differently (essentially the same, but only uses the coefficients corresponding to the active points).

\subsubsection{Washout of $G_2$}

The washout of $G_2$ is equivalent to the following hypothesis:

\begin{hypo}
	$G_2 \subset \mathcal{J}$ and $G_1 \not \subset \mathcal{J}$
	\label{WashoutG2}
\end{hypo}

In the case that $x_i = 0, \, \forall i \in G_2$ then from equation \eqref{Eq4}.
\begin{align}
\dot{s_3} = 0 = -s_3D \\
\Rightarrow s_3 = 0;
\end{align}

Then by \eqref{s1s2s3} one obtains:

\begin{align}
s_{in} = s_1 + s_2
\end{align}

Then by replacing  \eqref{EqSome} and \eqref{EquilibriaFormula} in \eqref{Eq2}.
\begin{align}
\label{S1NOBWashout}  (s_{in}-s_1)-1_{(n-m)\times 1}^\top
\diag(k)(A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1}) = 0 
\end{align}

Notice \eqref{S1NOBWashout} does not contain terms with $s_2$. By reordering the terms \eqref{S1NOBWashout} one finds a two degree polynomial of $s_1$:

\begin{align}
a'_2 s_1^2 + a'_1s_1 + a'_0 = 0
\end{align}

Where $a'_i$ can be found in the appendix.

\subsubsection{Washout}	
Notice that $x_i = 0$ for every $i \in \{1,\dots, n_1\}$ implies the washout equilibria (no OTU surviving):

From equation \eqref{Eq2}, one gets 
\begin{align*} s_{in} = s_1 \end{align*}
then \eqref{s1s2s3} implies \begin{align*}s_2 &= s_3 = 0 \\ \Rightarrow \dot{x_i} &= -Dx_i \quad \forall i = n_1+1, \dots, n\end{align*}   
which has an equilibria at $x_i = 0$.

All the former discussion leads to a potential number of   $ (2^{n_1}-1)\cdot (4\cdot(2^{n_2}-1)+2)+1 $  different equilibria. 
\begin{align} \overbrace{(2^{n_1}-1)}^{\substack{\text{nonempty subsets} \\ \text{of }G_1 }} (\underbrace{4}_{\substack{\text{possible} \\ \text{solutions of} \\ \text{equation \eqref{Poly4}}}} \cdot\overbrace{(2^{n_2}-1)}^{\substack{\text{nonempty subsets} \\ \text{of  }G_2}}+\underbrace{2}_{\substack{G_2 \\ \text{Washout}}})+\overbrace{1}^\text{Washout} 
\end{align}

\subsection{Local Stability and Operating Diagrams}
For checking local asymptotic stability of the equilibria points, the Jacobian of the system is provided and evaluated at each of these points. The resulting matrix's eigenvalues must have negative real part. A general formula for this Jacobian is presented in the appendix.

The former discussion proposes an algorithm for finding the positive stable equilibrium points

\begin{algorithm}[H]
	\KwData{$A \in M_{n\times n}(\R)$, $D, s_{in}, \bar{\mu}_i,K_i, k_i,  \in \R \, i\in [n], $, $n_1, n_2 \in \N$}
	\KwResult{Set PSP containing all Positive Stable Equilibrium Points}
	$PSP = \emptyset$ \\
	eig = Eigenvalues of $J(0,s_{in},0,0)$. \\
	\If{$Real(eig) < 0$}{$PSP = PSP \cup (0,s_{in},0,0)$} 
	\For{$A'$ square submatrix of $A$}{	Calculate $(x,s_1,s_2,s_3)$ from equations \eqref{Poly4}, \eqref{S2(S1)}, \eqref{s1s2s3}, and \eqref{xEquilibria} \\
		\If{$(x,s_1,s_2,s_3) \geq 0$}{eig = Eigenvalues of $J(x,s_1,s_2,s_3)$ \\
			\If{$Real(eig) < 0$}{$PSP = PSP \cup (x,s_1,s_2,s_3)$}}}
	\label{StabilityAnalysis}
\end{algorithm}

The algorithm implicitly assumes that every submatrix $A'$ is invertible.

Operating diagrams are an illustrative way of representing the long term behaviour of a reactor depending on operating parameters. Operating diagrams are created by running algorithm \eqref{StabilityAnalysis} for different pairs $(s_{in},D)$ \cite{lobry2017chemostat}. One then regroups all the pairs such that the set of non-active coordinates of the points of the set $PSP$ are the same and have the same stability properties. This regrouping represents graphically the coexistence of the different OTU in a $D-s_{in}$ plot. In particular for the biological process of nitrification, this allows to represent when partial nitrification (PN) or complete nitrification (CN) may arise \cite{KHIN2004519}. PN refers to the state when nitrite $s_2$, accumulates because the OTU of $G_2$ concentration is very low and thus almost no conversion from $s_2$  to $s_3$ takes place. On the contrary CN is when nitrate ($s_3$) accumulates because of the presence of OTU of $G_2$.

The operating diagram from one of the sets of the parameter identification from Dumont et al. is presented in figure \ref{OD Dumont} from a Bayesian estimation method. Two species per functional group were considered. One can see that coexistence in the same functional group is never attained. However when using a different set of parameters one can see a more complex situation where coexistence in the same functional group is possible as seen in figure \ref{OD modified}. In the latter case CN can be achieved at a more elevated dilution rats than in the former case. Parameters of cases 1 and 2 can be seen in table \ref{Parameters Dumont} and \ref{Parameters Modified}, respectively.

The model is also  able to overcome the  competitive exclusion principle \cite{lobry2017chemostat} (CEP) paradigm. The CEP roughly states that if two species are growing on the same limiting resource, and their growth law is monotonic in the substrate, then only one of them will survive in the long run. However reports on wastewater treatment plants have shown coexistence between species in nitrifying reactors has\cite{Wagner2002}.

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Parameters Dumont & $\mu_i\,[1/day]$ & $K_i\,[mg/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ & 0.828  & 0.147 & 3.85  \\ \hline
		$x_2\in G_1$ &0.828   & 0.147 &  3.85\\ \hline
		$x_3\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
		$x_4\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
	\end{tabular}
	\caption{Kinetic parameters of model \eqref{system} from \cite{Dumont2016}.}
	\label{Parameters Dumont}
\end{table}

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{Interactions_Parameters_Dumont}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{OD_parameters_Dumont}
	\end{subfigure}
	\caption{Left: interactions of the system. Right: operating diagram of Model \eqref{system}, with parameters in Table \ref{Parameters Dumont}. The different zones represent the surviving species in the steady state. PN takes place when neither $x_3$ nor $x_4$ are present. CN takes place if $x_3$ or $x_4$ are present.}
	\label{OD Dumont}
\end{figure}

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Modified Parameters  & $\mu_i\,[1/day]$ & $K_i\,[mg/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ &0.828  &0.147  & 5 \\ \hline
		$x_2\in G_1$ & 0.621 & 0.0147 & 3 \\ \hline
		$x_3\in G_2$ & 0.18 & 0.026 & 110 \\ \hline
		$x_4\in G_2$ & 0.135& 0.0026& 90 \\ \hline
	\end{tabular}
	\caption{Kinetic parameters of model \eqref{system} modified from \cite{Dumont2016}.}
	\label{Parameters Modified}
\end{table}
\begin{figure}[ht]
	\centering	
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{Interactions_Parameters_modifed}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{OD_parameters_modified}
	\end{subfigure}
	\caption{Left: interactions of the system. Right: operating diagram of Model \eqref{system}, with parameters in Table \ref{Parameters Modified}. The different zones represent the surviving species in the steady state. PN takes place when neither $x_3$ nor $x_4$ are present. CN takes place if $x_3$ or $x_4$ are present.}
	\label{OD modified}
\end{figure}
\clearpage

\subsection{Remarks}

The model here presented serves to illustrate that by considering a more complex growth rate that accounts for ecological interactions one might explain differences in reactors operating under similar conditions. It also shows a new mechanism by which the CEP no longer holds. However the quadratic increase of parameters with the number of OTU considered makes it difficult to accurately calibrate the model. The operating diagrams always show a single stable equilibrium point for each pair $(s_in,D)$, however no proof is offered to such conjecture. In the next section, a generalization of the former model, and a schema for calibrating a class of models is presented as well.

\section{Generalized approach for modelling interactions}
In the previous sections interactions were modelled as an affine function of the OTU concentration that multiplies a substrate dependent growth equation.

More generally the interaction function is defined. It represents how the growth rate of species $i$ is affected by the concentration of other species, $x$:

Given a vector $(v_1,\dots,v_n)^\top$ the interaction function $\I$ is noted as:
\begin{align}
\begin{array}{rc}
\I: \R_+^n \rightarrow & \R_+^n\\
& \\
v \rightarrow & \begin{pmatrix}
I_1(v) \\
I_2(v) \\ 
\vdots  \\
I_n(v)
\end{pmatrix}
\end{array}
\end{align}

	Let $f_i(s)$ be a bounded, positive, and continuous function of $s$ (e.g. Monod, Haldane). The growth equation of OTU $i$ becomes:
	\begin{align}
	\mu_i(s,x) = f_i(s)I_i(x)
	\end{align}
	Note $f(s) := (f_1(s),\dots,f_n(s))^\top$.
	\label{growthForm}

Since the growth of a single strain in batch experiments is driven by the substrate concentration, when no interactions are present one should recover expression $f_i(s)$. Therefore if there are no interactions then $I_i(x) = 1$. From this hypothesis, note that $ \lim \limits_{x \rightarrow 0} I_i(x) = 1$ since if there is minimal presence of OTU, interactions can not exist. Furthermore for this first study it is assumed that $I_i(\cdot)$ is a continuously differentiable function on $x$. For making explicit all of the former:

\begin{hypo}
	The interaction function $\I$ previously defined satisfies:
	\begin{enumerate}
		\item\begin{align}
		\I \left( (0,\dots,0)^\top \right) = 
		(1,\dots,1)^\top 
		\end{align}
		\item Let $\Omega$ be an open subset of $ \R^n $ then $\I \in C^1(\Omega)$.
	\end{enumerate} 
\end{hypo}

Note $J_I(x)$ the Jacobian matrix of function $\mathcal{I}$. then a first order approximation of $\I(\cdot)$ centred at $\bar{x}$ gives: $\I(x) = \I(\bar{x}) + J_I(\bar{x})(x-\bar{x}) + o(\Vert x- \bar{x} \Vert)$. I $\bar{x}= 0$, one recovers the growth expression from the previous section and $J_I(\bar{0})$ can be seen as the interaction matrix from model \eqref{system}.

\subsection{Reconstructing the Interaction Function}

Suppose that the functions $f_i(s)$, and the yields $y_i$ are well-known. By using experimental measurements of $x$, represented by $z(t)$, the objective is to reconstruct function $I(x)$. For doing so, the terms $I_i(x)$ are replaced by a control $u_i(t)$  and a state feedback is obtained by solving a nonlinear optimal tracking problem.

Consider the observable system \eqref{Controlsystem} ($y(t)$ being the observer) where $g(\cdot)$ is a known function.

\begin{align} 
\label{Controlsystem}
\begin{array}{cl}
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_1\\
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_2\\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D-\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)u_i(t) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)u_i(t)	-\sum\limits_{i \in G_2}\frac{1}{y_i}f_i(s)u_i(t) x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i \in G_2}\frac{1}{y_i}f_i(s)u_i(t) x_i \\
y(t) & = g(x,s)
\end{array}
\end{align}

For this model $g(x,s) = x$, since we are observing measurements coming from genetic sequencing. Consider the weighted norms defined by positive definite matrices $Q$ and $R$, represented by $\Vert \cdot \Vert_Q$ and $\Vert \cdot \Vert_R$, respectively. The optimal tracking problem is defined as:

\begin{align}
\label{Optimal Control} \begin{array}{cc} \min &  \int \limits_{0}^{T} \Vert g(x,s) - z \Vert_Q + \Vert(u -\vec{1}) \Vert_R dt\\
s.t.& 
(x,s_1,s_2,s_3)\mbox{ solution of \eqref{Controlsystem}} \\
&u_i(t) \in [0,1]
\end{array}	
\end{align}

The control $u(t)$ is intended to drive the system to be near a desired output $z(t)$, which in this context are the measurements of the concentrations of OTU and metabolites. The term $\lambda \Vert(u -\vec{1}) \Vert_R$, was added for two reasons:

First because the interest is testing the idea that interactions could be driving the system. Therefore adding a penalization in the objective function for each control to remain near $1$ can be seen as an attempt to explain data without any interaction.

Second, to force a regularized control. Otherwise note that $u$ is linear in \eqref{Controlsystem}, therefore if the integral cost does not have a non-linear expression of $u$ the optimal control will be of a bang-bang type with possibly singular arcs. Since the objective is to find a differentiable expression of $I(x)$ the addition of the regularization term is deemed necessary.

The problem of approximating the solution of the system to a desired reference ($z$ in this case) is called the optimal tracking problem. For solving such a problem the approach developed by Cimen \textit{et al.} \cite{Cimen2004} was adapted to our problem. The method proposed involves the resolution of Approximating Sequences of Ricatti Equations (ASRE).

The method consists on iteratively calculating trajectories of system \eqref{Controlsystem} with a certain control and feeding a non-autonomous Ricatti differential equation with the resulting trajectory. Then a control law that uses the Ricatti equation is proposed and a new trajectory is calculated. The iteration is stopped when a convergence in the output or in the control is observed.

Two points of departure from the original method are made clear, details can be seen in the appendix: first the tracking problem does not consider a constrained control, so one should consider a reformulation as proposed by the same author in other works \cite{Cimen2008}. Nevertheless, the methods were directly used, and an explicit constraint in the synthesis of the control was added for the control to remain within the desired bounds.

First the tracking problem does not consider a constrained control, so one should consider a reformulation as proposed by the same author in other works \cite{Cimen2008}. Nevertheless, the methods in \cite{Cimen2004} were directly used, and an explicit constraint in the synthesis of the control was added for the control to remain within the desired bounds (see equation \eqref{control synthesis}). Even though this is probably suboptimal it worked for our purposes.

Second in their problem they used an approximation of their original dynamics for calculating the trajectories and subsequently proved that such linearisation converged to the original dynamics. In the case here presented the linearisation of the dynamics was not necessary.
Second, when implementing equation (12) of their article \cite{Cimen2008} a linearised dynamic is not used. Instead the original dynamic with the control given by equation (9) of their article is used. This modification of the original method Second in their problem they used an approximation of their original dynamics for calculating the trajectories and subsequently proved that such linearisation converged to the original dynamics. In the case here presented the linearisation of the dynamics was not necessary.

The change of variable $u_i(t) = v_i(t) + 1$ is used for technical reasons explained in the appendix section. This in turn implies $v_i(t) \in [-1,0]$.

The feedback control $v(t)$, will be of the form $v(t) = -R^{-1}\tilde{B}^\top(t)\left(\tilde{P}(t)x(t)-s_f(t)\right)$. Where matrix $P(t)$ and vector $s_f(t)$ solve differential equations. In the appendix one may prove that because of the structure of the system one only needs to calculate $2n$ differential equations for the synthesis of the control, instead of $(n+3)^2 + (n+3)$ that would imply the direct application of the method. Which renders the method- at least theoretically- scalable.

Finally for the parametric identification of a particular class of interaction functions the following static optimization problem is proposed. Suppose $u(t)$ is the optimal control of problem \eqref{Optimal Control}, and $x(t)$ is the associated trajectory. Suppose a set of parameters $\theta \in \Theta$, such that the interaction function has a explicit form that depends on this parameters $I_{\theta}(x)$ ($\theta$ would represent matrix $A$ from section 2 for example). Then one solves the following optimization problem:
\begin{align}
\label{Parameter Identification} \min & \quad \Vert I_{\theta}(x(\cdot)) - u(\cdot) \Vert_{L^2([0,T])} \\
s.t. &  \quad \theta \in \Theta \notag \\
& \quad t \in [0,T] \notag
\end{align}

In practice $t \in [0,T]$ from \eqref{Parameter Identification} is replaced by considering the time steps given by the differential equation solver.
\subsection{Proof of concept}

For compensating the lack of a theoretical optimal control, the approach presented in here was first tested against synthetic data. The first test consisted in simulating system \eqref{Controlsystem} with $I(x) = 1 + Ax$, for $n_1 = 10$ and $n_2 = 5$.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_Biomass_Iter_300}
	\caption{Total biomass. Dotted line: simulated data. Continuous line: the tracking procedure results.}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_AOB_Iter_300}
	\caption{AOB biomass. Dotted line: simulated data. Continuous line: the tracking procedure results.}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_NOB_Iter_300}
	\caption{NOB biomass. Dotted line: simulated data. Continuous line: the tracking procedure results.}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_metabolites_Iter_300}
	\caption{Metabolites. Dotted line: simulated data. Continuous line: the tracking procedure results.}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_Control_AOB_Iter_300}
	\caption{Control for the AOB population.}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{Synthetic_data_no_noise//061119_Try1_Control_NOB_Iter_300}
	\caption{Control for the NOB population.}
\end{figure}

Finally in the parametric identification one can compare the original matrix, with the identified one.

The procedure is repeated but with noise on the simulated trajectory.

Again the method follows the tendencies from data.

And the reconstructed matrix

\clearpage
\section{Application}

The choice of matrices $Q$ and $R$ are always of discussion in a quadratic regulator. $Q$ represents the covariance matrix of measurements, while the interpretation of $R$ is not clear. In the simulations here presented $Q = I_n$ and $R = 0.0001I_n$, found by trial and error in order for the control to fit accurately the species abundance. The method stopped after 110 iterations where each iteration took approximately 600 seconds of computing time in a computer equipped with 8gb of RAM memory and Intel core i3-7100U CPU 2,40 GHz.

The tracking problem was applied to data coming from a nitrification process; the experimental conditions are described in \cite{dumont2008observers}. For exploring the hypothesis of interactions as drivers of bioreactors performance environmental conditions should be kept as constant as possible. Therefore only data from day $183$ onwards was used because a change in the operating temperature happened at that point, which is known to have an effect on the kinetics.  For choosing which species belong in which functional group, the procedure described in \cite{Ugalde-Salas2019} was used: from day $183$ to day $230$: 17 OTU were identified in the $G_1$ group (known as Ammonium oxidizing bacteria AOB) and 8 in the $G_2$ functional group (known as Nitrite oxidizing bacteria NOB). The system state dimension becomes 28.

The knowledge of functions $f_i(s)$ was based on a study of nitrification's kinetic parameters\cite{Wiesmann1994}. Particularly given the system's ammonium and nitrate concentration a Monod function (eq \eqref{monod}) was used for $G_1$ and $G_2$. 
\begin{align}
\label{monod} f_i(s) &= \bar{\mu}_2 \dfrac{s_1}{K_1 +s_1} \, \forall i \in G_1 \\
\notag  f_i(s) &= \bar{\mu}_2 \dfrac{s_2}{K_2 +s_2 } \, \forall i \in G_2
\end{align}

For the reader to gain understanding of the situation, a simulation of the system using the experiments operating parameters ($D$ and $s_{in}$) is presented without control (i.e. $v(t)= 0$) in  Figure \ref{NoControlSimulation}. $s_3$ accumulates all along the trajectory, but when compared to data it is clear that $s_3$ stops accumulating after a while.

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{Prediction_No_Control_metabolites}
	\caption{Simulation when $u=1$. }
	\label{NoControlSimulation}
\end{figure}

When applying the presented method one obtains the simulation that can be seen in figure \ref{ControlSimulation}. The method captures the tendencies in the measured substrates, but not their absolute values. The simulation of each functional $G_1$ (AOB), and $G_2$ (NOB) can be seen in figures \ref{AOBSimulation} and \ref{NOB Simulation}, respectively. Note that the fit of $G_1$ seems better than the fit of $G_2$, which can be explained by the 1 order of magnitude difference of abundance between the members of the group.

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{260919_Try1_metabolites_Iter_110}
	\caption{Simulation applying the control strategy.}
	\label{ControlSimulation}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{260919_Try1_AOB_Iter_110}
	\caption{Abundance of $G_1$ (AOB) simulation applying the control strategy.}
	\label{AOBSimulation}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{260919_Try1_NOB_Iter_110}
	\caption{Abundance of $G_2$ (NOB) simulation applying the control strategy.}
	\label{NOB Simulation}
\end{figure}

One can observe that in figure \ref{ControlSimulation} the predicted $s_3$ around day 240 is much higher than the measurement. If one looks at figure \ref{NOB Simulation} it suggests attributing this extra production of $s_3$ around day 240 to the growth of OTU 24. What would have happened if OTU 24 had been classified in the other functional group? One would have seen a decrease in $s_1$, an increase in $s_2$, and a decrease in $s_3$ which could have better fitted the measurements. Thus suggesting that OTU 24 was misclassified from the method presented in \cite{Ugalde-Salas2019}.

In figures \ref{AOB_control} and \ref{NOB_control} it can be seen the obtained controls for each functional group. Both controls saturate at certain points, but in general terms they vary throughout the dynamic.

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{260919_Try1_Control_AOB_Iter_110}
	\caption{Control obtained for OTU in $G_1$ (AOB).}
	\label{AOB_control}
	\includegraphics[width= 0.4\textwidth]{260919_Try1_Control_NOB_Iter_110}
	\caption{Control obtained for OTU in $G_2$ (NOB).}
	\label{NOB_control}
\end{figure}

Recalling that $u(t)$ is supposed to represent $I(x(t))$ (the interactions terms), one could try to fit different forms of $I(\cdot)$ for example $I(x) = 1 + Ax$, where  $A$ is a real valued square matrix. In figure \ref{LV_fit} the simulation of the system after the identification of matrix A is presented, suggesting that an affine interaction function is not able to capture the behaviour of the system.

\begin{figure}[h]
	\centering
	\includegraphics[width= 0.4\textwidth]{260919_Try1_metabolites_Iter_110_LV}
	\caption{Simulation obtained for the interaction function $I(x) = 1 + Ax$, where $A$ was fitted with the obtained controls from figures \ref{AOB_control} and \ref{NOB_control}.}
	\label{LV_fit}
\end{figure}
\clearpage
\section{Conclusion}

This article dealt with the problem of how to include genetic data in the prediction of performance of bioreactors, a previous model which deals with such an issue was revisited. The former model suffers from poor scalability, and therefore its application is limited. Inspired by their model a novel approach was here developed in order to systematically integrate the available genetic data. The approach is fairly straightforward for chemostat systems because it consists in multiplying a control to the growth expression of each OTU to be tracked. Nevertheless a method for previous identification of the microbial community must be used beforehand.
All the codes can be downloaded from Github.

\section{Acknowledgements}
 Alain Rapaport, CMM, Thermomic

\section{Appendix}

\textbf{Proof of lemma \ref{l1}}
\begin{proof}
	Define 
	\begin{align}
	z_1 &: = \sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1  \\
	z_2 &: = \sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 + s_2 \\
	z_3 &: = s_1 + s_2 + s_3
	\end{align}

	Computing $\dot{z_{1}}$ one gets:
	\begin{align*}
	\dot{z_{1}} &=\sum \limits_{i \in G_1} \frac{1}{y_i}\dot{x_i} + \dot{s_1} \\
	& = \sum \limits_{i \in G_1} \frac{1}{y_i}\left(\mu_i(s_1,x) -D \right)x_i + (s_{in}-s_1)D-\sum\limits_{i \in G_1}\frac{1}{y_i}\mu_i(s_1,x) x_i   \\
	&= 	D\left(-\sum \limits_{i \in G_1} \frac{1}{y_i}x_i - s_1 + s_{in}\right)  \\
	& = D(s_{in} - z_1)
	\end{align*}

	Define $\bar{s} =\max\{s_{in}(t) | t\geq 0\}$ and consider the differential equation:

	\begin{align}
	\label{compareEquation}
	\begin{array}{l}
	\dot{w} = D(\bar{s} - w) \\
	w(0) = \sum \limits_{i \in G_1} \frac{1}{y_i}x_i(0) + s_1(0)
	\end{array}
	\end{align}

	If there is a time interval $H$ such that $t^*\in H  \Rightarrow D(t^*) = 0$, then $z_1$ is constant and therefore bounded by  the value of the solution in $z(t^*)$. 
	If $D(t) > 0$, then $w = \bar{s}$ is a stable asymptotic equilibrium. Define $M_1 :=  \max\{w(0),\bar{s}\}$. If $ w(0) > \bar{s}$ then by the Picard-Lindeloff theorem $ \forall t \; \dot{w}(t) < 0 $, otherwise it would cross the solution of the initial value problem \eqref{compareEquation} with starting point $\bar{s}$ , therefore $w(t) \geq \bar{s}$. The same reasoning may be applied if $ w(0) < \bar{s}$. One concludes that $ w(t) \leq M_1 $.

	Consider  
	\begin{align}
	\label{z1} \begin{array}{l}
	\dot{z_1} =  D(s_{in} - z_1)\\
	z_1(0) = \sum \limits_{i \in G_1} \frac{1}{y_i}x_i(0) + s_1(0)
	\end{array} 
	\end{align}
	From a comparison lemma (see chapter 3 \cite{Khalil1996}), the solution of \eqref{z1} is bounded by $w$ and therefore $z_1(t) \leq M_1$.

	By defining $M_2 = \max\left \{ \displaystyle \sum \limits_{i=n_1 +1 }^{n} \frac{1}{y_i}x_i(0) + s_1(0) +s_2(0), \bar{s} \right\}$, and $M_3 =\max\left \{ s_1(0) +s_2(0) + s_3(0), \bar{s} \right\} $ and noting that $z_2(t)$ and $z_3(t)$ satisfies the same differential equations as $z_1$ the above reasoning may be applied and one has the desired bounds.
\end{proof}

\textbf{Proof of theorem \ref{theoWellPosedness}}

\begin{proof}
	If any coordinate of the solution becomes zero, then it's derivative is either zero, or positive. The bound will follow from Lemma \ref{l1}.

	Note that for any $i \in \{1,\dots, n\} $ if $x_i = 0$ then $\dot{x_{i}} = 0$, therefore all solutions of system \eqref{system} with initial condition $x_i(t_0) = 0$ remain in these planes. By the Picard Lindeloff theorem a solution starting in $int(\Omega)$ cannot cross these planes, therefore $x_i(t)\geq 0$ for $t \geq 0$.

	If $s_1 = 0$, then $\dot{s_1} = Ds_{in} > 0$. Therefore $s_1(t) \geq 0$.

	If $s_2 = 0$ then note first that by adding both inequalities from Lemma \eqref{l1} and since, $s_1,x_i \geq 0$ one has:
	\begin{align*}
	\sum \limits_{i \in G_1 \cup G_2}  \frac{1}{y_i} x_i + 2s_1 + s_2 \leq M_1 + M_2 \\
	\Rightarrow \sum \limits_{i = 1}^{n}  \frac{1}{y_i} x_i + 2s_1 \leq M_1 + M_2 \\
	\Rightarrow \sum \limits_{i = 1}^{n}  \frac{1}{y_i} x_i \leq M_1 + M_2\\ 
	\Rightarrow \bar{k} \sum  \limits_{i = 1}^{n}   x_i  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \sum  \limits_{i = 1}^{n}   \vert x_i \vert  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \Vert x \Vert_1  \leq M_1 + M_2  \\
	\Rightarrow   \bar{k} \Vert x \Vert_{\infty} \leq M_1 + M_2
	\end{align*}

	Where $\bar{k} = \min \left\{ \frac{1}{y_i}  : \; i\in [n] \right \}$.

	Define $M := \dfrac{\bar{k}}{M_1 + M_2}$, and let $A$ be a matrix such that $\Vert A \Vert_{\infty} \leq M$. Then since $\Vert Ax \Vert_{\infty} \leq 1 \Rightarrow \forall i=1,\dots,n \; (1 + A_{i\bullet}x) \geq 0 $, and $\Vert Ax \Vert_{\infty} \leq \Vert A \Vert_{\infty} \Vert x\Vert_{\infty} \leq M \frac{1}{M}$.
	One has proven that:
	\begin{align}
	\label{Bound1} \forall i=1,\dots,n \;  \;  (1 + A_{i\bullet}x) \geq 0 
	\end{align}

	Therefore $\dot{s_2} = \displaystyle \sum \limits_{i \in G_1 } \frac{1}{y_i}\mu_i(s_1,x)x_i =  \sum \limits_{i \in G_1 } \frac{1}{y_i}\bar{\mu_i}f_i(s)(1+A_{i \bullet }x)x_i \geq 0$. Therefore $s_2(t) \geq 0$.

	Note that since one has proven that $s_2 \geq 0$, bound \eqref{Bound1} is valid for any time, and not only when $s_2 = 0$. 
	If $s_3 = 0$ then $\dot{s_3} = \sum \limits_{i \in G_2 }^{n}\frac{1}{y_i}\mu_i(s_2,x)\geq 0$. Therefore $s_3 \geq 0$.

	For the boundedness it suffices to notice from Lemma \ref{l1} that the sum of positive elements is bounded, therefore each element is bounded.
\end{proof}

\subsection{Properties of diag operator}

Notice that for two vectors $v_1$ and $v_2$ the following holds:
\begin{align}
\diag(v_1)\diag(v_2) = \diag(v_2)\diag(v_1) 
\end{align}

Suppose $v \in \R^n$ such that $v_i \neq 0 \forall i \in [n]$. Define $v^{-1}:= \left( \dfrac{1}{v_1},\dots,\dfrac{1}{v_n} \right)$ then the inverse of $\diag(v)$ exists and is $\diag(v^{-1})$:

\begin{align}
\diag(v)^{-1} = \diag(v^{-1}) 
\end{align}

\subsection{Inverse of A and products}
Notice that $\begin{bmatrix}
A_1 \\ A_2
\end{bmatrix} A^{-1} = \begin{bmatrix}
A_1A^{-1} \\ A_2A^{-1}
\end{bmatrix}=I $.

Then $A_1A^{-1}= \begin{bmatrix}
I_{n_1}& 0_{n_2 \times 1} 
\end{bmatrix}$ and $A_2A^{-1}= \begin{bmatrix}
0_{n_1 \times 1} & I_{n_2}
\end{bmatrix}$.

For notation purposes it is useful to define: \begin{align}
B := \diag(k) A^{-1}.
\end{align}

\subsection{Deduction of  \eqref{s1Equilibria}}

Replacing \eqref{xEquilibria} in \eqref{Eq2} yields :
\begin{align}
(s_{in}-s_1)D- \begin{bmatrix}
1_{n_1 \times 1}+A_1A^{-1}(\diag(f(s))^{-1}D_{n_1 \times 1}-1_{n_1 \times 1}) \\0_{n_2 \times 1}
\end{bmatrix}^\top \diag(k)\diag(f(s))A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1})=0 \\
(s_{in}-s_1)D-	\begin{bmatrix}
1_{n_1 \times 1}+\begin{bmatrix}
I_{n_1} & 0_{n_2 \times 1} 
\end{bmatrix}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) \\0_{n_2\times 1}
\end{bmatrix}^\top \diag(k)\diag(f(s))A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) = 0 \\
(s_{in}-s_1)D-	\begin{bmatrix}
D_{n_1 \times 1}\\0_{n_2 \times 1}
\end{bmatrix}^T\diag(f(s))^{-1}\diag(f(s)) \diag(k) A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) = 0  \\
(s_{in}-s_1)D-\begin{bmatrix}
D_{n_1 \times 1} \\0_{n_2\times 1}
\end{bmatrix}^\top \diag(k)A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n_1 \times 1}) = 0 \\
(s_{in}-s_1)- \begin{bmatrix}
1_{n_1 \times 1}\\0_{n_2 \times 1}
\end{bmatrix}^\top B (\diag(f(s))^{-1}D_{n \times 1} - 1_{n \times 1}) = 0 
\end{align}

\subsection{Deduction of \eqref{s2Equilibria}}
Analogous as before, replacing \eqref{xEquilibria} in \eqref{Eq3} yields :
\begin{align}
-s_2D+
\begin{bmatrix}
(\diag(\mu_A)^{-1} \vec{D}) \\ -(\diag(\mu_B)^{-1} \vec{D})
\end{bmatrix}^\top \diag(\mu)\diag(y)^{-1} A^{-1}(\diag(\mu)^{-1}\vec{D}-\vec{1}) = 0 \\
-s_2D+
\begin{bmatrix}
\vec{D} \\ - \vec{D}
\end{bmatrix}^\top \diag(y)^{-1} A^{-1}(\diag(\mu)^{-1}\vec{D}-\vec{1}) = 0\\
-s_2+
\begin{bmatrix}
\vec{1} \\ - \vec{1}
\end{bmatrix}^\top B(\diag(\mu)^{-1}\vec{D}-\vec{1}) = 0
\end{align}
\begin{align} 
-s_2D+ \begin{bmatrix}
1_{n_1 \times 1}+A_1A^{-1}(\diag(f(s))^{-1}D_{n_1 \times 1}-1_{n_1 \times 1}) \\-(1_{n_1 \times 1}+A_2A^{-1}(\diag(f(s))^{-1}D_{n_1 \times 1}-1_{n_1 \times 1})
\end{bmatrix}^\top \diag(k)\diag(f(s))A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) = 0 \\
-s_2D+	\begin{bmatrix}
1_{n_1 \times 1}+\begin{bmatrix}
I_{n_1} & 0_{n_2 \times 1} 
\end{bmatrix}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) \\-(1_{n_1 \times 1}+\begin{bmatrix}
0_{n_1 \times 1} & I_{n_2} 
\end{bmatrix}(\diag(f(s))^{-1}D_{n_1 \times 1}-1_{n_1 \times 1})
\end{bmatrix}^\top \diag(k)\diag(f(s))A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) = 0 \\
-s_2D +	\begin{bmatrix}
D_{n_1 \times 1}\\-D_{n_2 \times 1}
\end{bmatrix}^T\diag(f(s))^{-1}\diag(f(s)) \diag(k) A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n \times 1}) = 0  \\
-s_2D+\begin{bmatrix}
D_{n_1 \times 1} \\-D_{n_2\times 1}
\end{bmatrix}^\top \diag(k)A^{-1}(\diag(f(s))^{-1}D_{n \times 1}-1_{n_1 \times 1}) = 0 \\
-s_2 + \begin{bmatrix}
1_{n_1 \times 1}\\1_{n_2 \times 1}
\end{bmatrix}^\top B (\diag(f(s))^{-1}D_{n \times 1} - 1_{n \times 1}) = 0 
\end{align}
\subsection{Deduction of \eqref{S2(S1)}}

Note that 
\begin{align}  (\diag(f(s))^{-1}D_{n\times 1}-1_{n\times1})  & = \begin{array}{cc} 
\begin{bmatrix}
\left(D\frac{K_1+s_1}{\bar{\mu}_{1}s_1}-1 \right) \\ \vdots \\ \left(D\frac{K_n+s_2}{\bar{\mu}_{n}s_2}-1\right) 
\end{bmatrix} & \hspace{-0.5cm}\begin{array}{c} \left\} n_1\right. \\ \\ \left\} n_2\right.\end{array}
\end{array} \\
\label{invMu} 
& = \begin{bmatrix}
\dfrac{DK_1+(D-\bar{\mu}_1)s_1}{\bar{\mu}_{1}s_1} \\ \vdots \\ \dfrac{DK_n+(D-\bar{\mu}_n)s_2}{\bar{\mu}_{n}s_2} 
\end{bmatrix}
\end{align}

Replacing \eqref{invMu} in \eqref{s1Equilibria}.

\begin{align}
(s_{in}-s_1)-	\begin{bmatrix}
1_{n_1 \times 1} \\0_{n_2\times 1}
\end{bmatrix}^\top  B\begin{bmatrix}
\dfrac{DK_1+(D-\bar{\mu}_1)s_1}{\bar{\mu}_{1}s_1} \\ \vdots \\ \dfrac{DK_n+(D-\bar{\mu}_n)s_2}{\bar{\mu}_{n}s_2} 
\end{bmatrix} = 0 \\
(s_{in}-s_1) - \sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}s_1} +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}s_2} \right) = 0 \quad / \cdot s_1s_2 \\
(s_{in}-s_1)s_1s_2 - \sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}}s_1 \right) = 0 \\
(s_{in}-s_1)s_1s_2 - \sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}s_2s_1 + \sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_1 \right) = 0  \\
s_2\left((s_{in}-s_1)s_1 - \sum \limits_{i=1}^{n_1}  \left( \sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}} + \sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}s_1 \right)  \right) =  \sum \limits_{i=1}^{n_1}\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_1 \\
s_2\left(-s_{1}^2  +  \left( s_{in}\underbrace{- \sum \limits_{i=1}^{n_1} \sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}}_{\beta_1}\right) s_1 \underbrace{ - \sum \limits_{i=1}^{n_1}  \sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}}_{\beta_2}  \right) = \underbrace{\sum \limits_{i=1}^{n_1}\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}}_{\beta_3}s_1 \\
s_2\left(-s_{1}^2  +  \left( s_{in}+\beta_1 \right) s_1 +\beta_2  \right) = \beta_3 s_1
\end{align}

From the last expression it should be clear that formula \eqref{S2(S1)} holds, it is just a matter of identifying the coefficients.
\begin{itemize}
	\item $\beta_1 =  - \sum \limits_{i=1}^{n_1} \sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}$
	\item $\beta_2 =-\sum \limits_{i=1}^{n_1}  \sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}$
	\item $\beta_3 = \sum \limits_{i=1}^{n_1}\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}} $
	\item $c_1 = \dfrac{-1}{\beta_3}$
	\item $c_2 =  \dfrac{s_{in} + \beta_1}{\beta_3}$
	\item $c_3 = \dfrac{\beta_2}{\beta_3}$
\end{itemize}

\subsection{Deduction of \eqref{Poly4}}

Again, replacing \eqref{invMu} in \eqref{s2Equilibria}.
\begin{align}
-s_2+
\begin{bmatrix}
1_{n_1 \times 1}\\ - 1_{n_2 \times 1}
\end{bmatrix}^\top B\begin{bmatrix}
\dfrac{DK_1+(D-\bar{\mu}_1)s_1}{\bar{\mu}_{1}s_1} \\ \vdots \\ \dfrac{DK_n+(D-\bar{\mu}_n)s_2}{\bar{\mu}_{n}s_2} 
\end{bmatrix} = 0 \\
\begin{array}{c} \displaystyle  -s_2 + \sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}s_1} +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}s_2} \right) \\
\displaystyle - \sum \limits_{i=n_1+1}^{n}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}s_1} +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}s_2} \right)= 0
\end{array} \quad / \cdot s_1s_2 \\
\begin{array}{c} \displaystyle  -s_1s_2^2 +\sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}}s_1 \right) \\
\displaystyle -  \sum \limits_{i=n_1+1}^{n}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_2}{\bar{\mu}_{j}}s_1 \right)= 0
\end{array} \\
\begin{array}{c} \displaystyle  -s_1s_2^2 + \sum \limits_{i=1}^{n_1}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}s_2s_1 + \sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_1\right) \\
\displaystyle -  \sum \limits_{i=n_1+1}^{n}\left(\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_2 +\sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}s_2s_1 + \sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}s_1 \right)= 0
\end{array}\\
\begin{array}{c} \displaystyle  -s_1s_2^2  - \beta_2s_2 -\beta_1s_1s_2 + \beta_3s_1 
\underbrace{- \sum \limits_{i=n_1+1}^{n}\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}}_{\beta_4}s_2 \underbrace{-\sum \limits_{i=n_1+1}^{n}\sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}}_{\beta_5}s_2s_1\underbrace{ -\sum \limits_{i=n_1+1}^{n}\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}}_{\beta_6}s_1 = 0
\end{array}\\
-s_1s_2^2  - \beta_2s_2 -\beta_1s_1s_2 + \beta_3s_1 + \beta_4s_2 +\beta_5s_1s_2 + \beta_6s_1  = 0 \\
\label{Prepoly4} (\beta_4-\beta_2)s_2 + (\beta_3+\beta_6)s_1 + (\beta_5 - \beta_1)s_1s_2- s_1s_2^2 = 0 
\end{align}
\begin{itemize}
	\item  $\displaystyle  \beta_4 = - \sum \limits_{i=n_1+1}^{n}\sum \limits_{j = 1}^{n_1} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}$
	\item $\displaystyle  \beta_5 = -\sum \limits_{i=n_1+1}^{n}\sum \limits_{j = 1}^{n} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}$
	\item $\displaystyle \beta_6 =  -\sum \limits_{i=n_1+1}^{n}\sum \limits_{j = n_1+1}^{n} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}$
\end{itemize}

Then replace \eqref{S2(S1)} in \eqref{Prepoly4}: 
\begin{align}
(\beta_4-\beta_2)\dfrac{s_1}{c_1s_1^2+c_2s_1+c_3} + (\beta_3+\beta_6)s_1 + (\beta_5 - \beta_1)s_1\dfrac{s_1}{c_1s_1^2+c_2s_1+c_3}- s_1\left(\frac{s_1}{c_1s_1^2+c_2s_1+c_3}\right)^2 = 0 \quad / \cdot \frac{1}{s_1} \\
(\beta_4-\beta_2)\dfrac{1}{c_1s_1^2+c_2s_1+c_3} + (\beta_3+\beta_6) + (\beta_5 - \beta_1)\dfrac{s_1}{c_1s_1^2+c_2s_1+c_3}- \left(\frac{s_1}{c_1s_1^2+c_2s_1+c_3}\right)^2 = 0  \quad / \cdot (c_1s_1^2+c_2s_1+c_3)^2 \\
(\beta_4-\beta_2)(c_1s_1^2+c_2s_1+c_3)+ (\beta_3+\beta_6)(c_1s_1^2+c_2s_1+c_3)^2 + (\beta_5 - \beta_1)(c_1s_1^2+c_2s_1+c_3)s_1 - s_1^2 = 0   \\
\begin{array}{c}(\beta_4-\beta_2)(c_1s_1^2+c_2s_1+c_3)+ (\beta_3+\beta_6)(c_1^2s_1^4+c_2^2s_1^2+c_3^2 + 2 c_1c_2s_1^3 + 2c_2c_3s_1 + 2 c_1c_3s_1^2)\\ \\  + (\beta_5 - \beta_1)(c_1s_1^2+c_2s_1+c_3)s_1 - s_1^2 = 0 \end{array} \\
\begin{array}{c} \underbrace{(\beta_3+\beta_6)c_1^2}_{a_4} s_1^4 + (\underbrace{(\beta_3 + \beta_6)(2c_1c_2) +  (\beta_5 -\beta_1)c_1}_{a_3})s_1^3 + (\underbrace{(\beta_4 - \beta_2)c_1 +(\beta_3 + \beta_6)(c_2^2 + 2c_1c_3) + (\beta_5-\beta_1)c_2-1}_{a_2})s_1^2 \\ 
+(\underbrace{(\beta_4 -\beta_2)c_2 + (\beta_3 + \beta_6)2c_2c_3+(\beta_5-\beta_1)c_3}_{a_1})s_1 + \underbrace{(\beta_4-\beta_2)c_3 + (\beta_3 + \beta_6)c_3^2}_{a_0} = 0
\end{array}
\end{align}

Which yields \eqref{Poly4}.
\begin{itemize}
	\item $a_4 = (\beta_3+\beta_6)c_1^2$ 
	\item $a_3 = (\beta_3 + \beta_6)(2c_1c_2) +  (\beta_5 -\beta_1)c_1$
	\item $a_2 = (\beta_4 - \beta_2)c_1 +(\beta_3 + \beta_6)(c_2^2 + 2c_1c_3) + (\beta_5-\beta_1)c_2-1$
	\item $a_1 = (\beta_4 -\beta_2)c_2 + (\beta_3 + \beta_6)2c_2c_3+(\beta_5-\beta_1)c_3$
	\item $a_0 = (\beta_4-\beta_2)c_3 + (\beta_3 + \beta_6)c_3^2$
\end{itemize}

\textbf{Note:} When calculating the equilibria with the extinction of some OTU under H \ref{noWashoutHyp}, coefficients are calculated with $B = \diag(k^{act})(A^{act})^{-1}$. The indexes are redefined to be taken only in $i,j \in [n] \setminus \mathcal{J}$ by formula \eqref{EquilibriaFormula}.

\subsection{Deduction of \eqref{S1NOBWashout}}
For auxiliary purposes $B:= \diag(k^{act})(A^{act})^{-1}$.

Under hypothesis \ref{WashoutG2}, the $j \in \mathcal{J}$ entries of vector $\diag(k) \diag(f(s))x^{eq} $ are zero. 
Note also that $(A_1x^{eq})_i = (A^{act}x^{act})_i, \, \forall i \in [n] \setminus \mathcal{J}$.

Then the cross product in equation \eqref{Eq2} can be reduced:
\begin{align} 
\begin{bmatrix}
1_{n_1\times 1}+A_1x^{eq} \\0
\end{bmatrix}^\top \diag(k) \diag(f(s))x^{eq} = [1_{(n-m)\times 1} - A^{act}x^{act}]^\top \diag(k^{act}) \diag(f^{act}(s))x^{act} 
\end{align} 
Therefore replacing \eqref{EquilibriaFormula} and \eqref{EqSome} in \eqref{Eq2}.

\begin{align}
0 &= (s_{in}-s_1)D- [1_{(n-m)\times 1} - A^{act}x^{act}]^\top \diag(k^{act}) \diag(f^{act}(s))x^{act}\\
&= (s_{in}-s_1)D- [1_{(n-m)\times 1} - A^{act}(A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1})]^\top \dots \\ & \diag(k^{act}) \diag(f^{act}(s))(A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1})\\
&  =  (s_{in}-s_1)-1_{(n-m)\times 1}^\top
B(\diag(f^{act}(s))^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1})
\end{align}

$s_2$ does not appear in the previous equation. Then as before (using the numeration of active indexes):
\begin{align}
\diag(\mu^{act})^{-1}D_{(n-m)\times 1} - 1_{(n-m)\times 1} = \begin{bmatrix}
\dfrac{DK_1+(D-\bar{\mu}_1)s_1}{\bar{\mu}_{1}s_1} \\ \vdots \\ \dfrac{DK_{n-m}+(D-\bar{\mu}_{n-m})s_1}{\bar{\mu}_{n-m}s_1}
\end{bmatrix}
\end{align}

Therefore:

\begin{align}
(s_{in}-s_1)- 1_{(n-m)\times 1}^\top 
B\begin{bmatrix}
\dfrac{DK_1+(D-\bar{\mu}_1)s_1}{\bar{\mu}_{1}s_1} \\ \vdots \\ \dfrac{DK_{n-m}+(D-\bar{\mu}_{n-m})s_1}{\bar{\mu}_{n-m}s_1}
\end{bmatrix} = 0 \\
(s_{in}-s_1) -\frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j=1}^{n-m} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}s_1}  = 0 \quad /\cdot s_1 \\
(s_{in} - s_1)s_1   -\frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j=1}^{n-m} B_{ij}\dfrac{DK_j+(D-\bar{\mu}_j)s_1}{\bar{\mu}_{j}}  = 0 \\
-s_1^2 + \left( s_{in}-\frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j=1}^{n-m} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}} \right)s_1 - \frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j=1}^{n-m} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}} =0
\end{align}

Identifying coefficients
\begin{itemize}
	\item $a'_2 = -1$
	\item $a'_1 = s_{in}-\frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j = 1}^{n-m} B_{ij}\dfrac{(D-\bar{\mu}_j)}{\bar{\mu}_{j}}$
	\item $a'_0 =- \frac{1}{y_A} \sum \limits_{i=1}^{n-m}\sum \limits_{j = 1}^{n-m} B_{ij}\dfrac{DK_j}{\bar{\mu}_{j}}$ 
\end{itemize}

\subsection{Jacobian of the system}
The jacobian is a square matrix of dimension $n\times 3$. The entries corresponding to the derivative of the first $n$ functions in the first $n$ variables can be calculated by the use of the chain rule. Define (for a fixed $s$):
\begin{align*} 
& \begin{array}{rc}
h_1:\mathcal{M}_{n \times n}(\R) \times \R^n \rightarrow & \R^n \\
(M,y) \rightarrow & My
\end{array}  \\
&\begin{array}{rl}
h_2: \R^n \rightarrow & \R^n \\
x \rightarrow & \diag(f(s))\left(1_{n\times 1}+ Ax\right)-D_{n\times 1}
\end{array} 	
\end{align*}

The first $n$ equations of system \eqref{system} may be written then as$\dot{x} =  h(x) = h_1(\diag(x), h_2(x))$, by the chain rule: 
\begin{align*} J_h(x) &= D_M h_1(\diag(x),h_2(x))\circ D_x \diag (x) + D_yh_1(\diag(x),h_2(x))\circ D_xh_2(x)  \\ &=
\diag(h_2(x))+\diag(x)\diag(f(s))A  \label{n2_entries}
\end{align*}

Expression \eqref{n2_entries} corresponds to the first $n\times n$ entries of the Jacobian of system \eqref{system}.

\begin{align*}
f_A' := \left( \dfrac{\partial f_1}{\partial s_1}, \dots, \dfrac{\partial f_{n_1}}{\partial s_1} \right)^\top = \left(\dfrac{\bar{\mu}_1K_1}{(K_1 + s_1)^2},\dots, \dfrac{\bar{\mu}_{n_1}K_{n_1}}{(K_{n_1} + s_1)^2} \right)^\top \in \R^{n_1} \\
f_B' := \left( \dfrac{\partial f_{n_1+1}}{\partial s_2}, \dots, \dfrac{\partial f_{n}}{\partial s_2} \right)^\top = \left( \dfrac{\bar{\mu}_{n_1+1}K_{n_1+1}}{(K_{n_1+1} + s_2)^2},\dots, \dfrac{\bar{\mu}_{n}K_{n}}{(K_{n} + s_2)^2} \right)^\top \in \R^{n_2}
\end{align*}

Then the Jacobian of the system may be expressed as:
$J(x,s_1,s_2,s_3) = \begin{bmatrix}
J_{11} & J_{12} & J_{13} & J_{14} \\
J_{21} & J_{22} & J_{23} & J_{24} \\
J_{31} & J_{32} & J_{33} & J_{34} \\
J_{41} & J_{42} & J_{43} & J_{44} 
\end{bmatrix} $

\begin{align}
J_{11} &=  \diag(\diag(f(s))(1_{n\times 1} +Ax)-D_{n\times 1})
+\diag(f(s))\diag(x)A \\
J_{12} &= \diag(x)\diag(1_{n \times 1}+Ax)\begin{bmatrix} f_A' \\ 0_{n_2 \times 1} \end{bmatrix}  \\
J_{13} &=\diag(x)\diag(1_{n\times 1}+Ax)\begin{bmatrix} 0_{n_1 \times 1} \\ f_B' \end{bmatrix} \\
J_{14} &= 0_{n\times 3}\\
J_{21} &= -\begin{bmatrix}
1_{n_1 \times 1} + A_1 x \\ 0_{n_2 \times 1} \end{bmatrix}^\top \diag(y)^{-1} \diag (f(s))  + (\diag(k)\diag(f(s))x)^\top\begin{bmatrix} A_1 \\ 0 \end{bmatrix}  \\
J_{22} &= -D  - \begin{bmatrix} 1_{n_1 \times 1} + A_1x \\ 0 \end{bmatrix}^\top \diag(k) \diag\left( \begin{bmatrix}\mu_A' \\ 0 \end{bmatrix}\right)x \\
J_{23} & = 0 \\
J_{24} &= 0 \\
J_{31} &= \begin{bmatrix} \left(1_{n_1 \times 1}+A_1x \right)\\ - \left(1_{n_2 \times 1} + A_2x \right)  \end{bmatrix}^\top \diag(k) \diag(f(s)) 
+  (\diag(k)\diag(f(s))x)^\top \begin{bmatrix}A_1 \\ -A_2\end{bmatrix} \\
J_{32} &= \begin{bmatrix}\left(1_{n_1 \times 1} + A_1x\right)\\ -\left(1_{n_2 \times 1} + A_2x\right)\end{bmatrix}^\top \diag(k) \diag\left( \begin{bmatrix}
f_A' \\ 0\end{bmatrix}\right)x \\
J_{33} &= -D + \begin{bmatrix} \left(1_{n_1 \times 1} + A_1x\right) \\ -\left(1_{n_2\times 1}+ A_2x\right) \end{bmatrix}^\top \diag(k)\diag\left(
\begin{bmatrix}
0 \\ f_B'
\end{bmatrix}\right)x  \\
J_{34} &= 0 \\
J_{41} &= \begin{bmatrix}0 \\ 1_{n_2 \times 1} + A_2x \end{bmatrix}^\top\diag(k) \diag(f(s)) 
+ (\diag(k)\diag(f(s))x)^\top\begin{bmatrix} 
0 \\ A_2
\end{bmatrix} \\
J_{42} &= 0 \\
J_{43} &= \begin{bmatrix} 0 \\(1_{n_2\times 1} + A_2x) \end{bmatrix}^\top \diag(k) \diag\left(\begin{bmatrix}
0 \\ f_B'
\end{bmatrix}\right) x \\
J_{44} &= -D
\end{align}

\subsection{Tracking Problem reformulation and details}

For applying the methods developed in \cite{Cimen2004}. Define the system state $X = (x,s)$. Make the change of variables $v_i = u_i - 1$ with $v = (v_1,\dots,v_n)$ are applied to system \eqref{Controlsystem}. The system may be rewritten then as:
\begin{align} 
\begin{array}{cl}
\dot{x_i} =& \left(f_i(s)(1+v_i(t)) -D \right)x_i \quad \forall i \in G_1\\
\dot{x_i} =& \left(f_i(s)(1+v_i(t)) -D \right)x_i \quad \forall i \in G_2\\
\dot{s_1} =& \displaystyle s_1\left(\frac{s_{in}}{s_1}-1\right)D-\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)(1+v_i(t)) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)(1+v_i(t))	-\sum\limits_{i \in G_2}\frac{1}{y_i}f_i(s)(1+v_i(t)) x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i \in G_2}^{n_1+n_2}\frac{1}{y_i}f_i(s)(1+v_i(t)) x_i \\
y(t) & = g(x,s)
\end{array}
\end{align}

Define $k_{G_1} = \left ( \frac{1}{y_1},\dots,\frac{1}{y_{nA}}\right )^\top \in \R^{n_1}$,  $k_{G_2} = \left(\frac{1}{y_{n_1+1}},\dots,\frac{1}{y_{n}}\right )^\top \in \R^{n_2}$, and

\begin{align}
A(X) & = \begin{bmatrix} 
\diag(f(s)- D_{n\times 1} ) & 0_{n \times 3}\\
\begin{matrix} 
f(s)^\top \begin{bmatrix} -\diag(k_{G_1}) & 0_{n_1 \times n_2} \end{bmatrix}^\top  \\
f(s)^\top \begin{bmatrix} \diag(k_{G_1}) & -\diag(k_{G_2}) \end{bmatrix}^\top   \\
f(s)^\top \begin{bmatrix} 0_{n_2 \times n_1} & \diag(k_{G_2}) \end{bmatrix}^\top  \end{matrix} & \begin{matrix} \left(\frac{s_{in}}{s_1}-1\right)D & 0 & 0 \\ 
0 &-D & 0 \\ 
0 & 0 &-D \end{matrix}
\end{bmatrix}  \\
B(X) & = \begin{bmatrix} \diag(f(s)\diag(x)  \\
\begin{matrix}
f(s)^\top \diag(x) \begin{bmatrix} -\diag(k_{G_1}) & 0_{n_1 \times n_2} \end{bmatrix}^\top \\
f(s)^\top \diag(x) \begin{bmatrix} \diag(k_{G_1}) & -\diag(k_{G_2}) \end{bmatrix}^\top  \\
f(s)^\top \diag(x) \begin{bmatrix} 0_{n_2 \times n_1} & \diag(k_{G_2}) \end{bmatrix}^\top  
\end{matrix}
\end{bmatrix} \\
C(X) = \begin{bmatrix}
I_n & 0_{n\times 3} \\
\end{bmatrix}
\end{align}

Then the system \eqref{Controlsystem} can be rewritten as:

\begin{align}
\dot{X} &= A(X)X + B(X)v \\
y &= C(X)X
\end{align}

$z(t)\in \R^n$ is the measured vector containing the OTU concentrations in time. The cost functional is given by
\begin{align}
J(v) = \left(z(t_f) - C(X) X(t_f)\right)^\top F\left(z(t_f) - C(X) X(t_f)\right) +  \int \limits_{t_0}^{t_f} \left(z(t) - C(X) X(t)\right)^\top Q \left(z(t) -C(X)X(t)\right) + v(t)^\top R v(t)
\end{align}

where $F,Q$ and $R$ are positive definite matrices. Particularly in all simulations $R = \lambda I_n$ for some $\lambda >0$, so all controls have the same weight. Since there is no interest in the final time, we define $F = 0$. $Q$ is a diagonal matrix.

Define the dynamic sequences for $i \in \N \cup \{0\}$, $\dot{X}^{[i]}$ as :

\begin{align}
\dot{X}^{[i]} &= A(X^{[i]})X^{[i]} + B(X^{[i]})v^{[i]} \quad i\in \N \\
y^{[i]} &= X^{[i]} \quad i\in \N \\
X^{[i]}(t_0) &= X_0 \quad i\in \N
\end{align}

And for $i = -1$ define $X^{[-1]}(t) = X_0$.

The control law is given by
\begin{align}
v^{[i]}(t)_j = \max \left\{ -1,\min\left\{0,\left( -R^{-1}B^\top\left(X^{[i-1]}(t)\right)\left(P^{[i]}(t)X^{[i]}(t)-s_f^{[i]}(t)\right)\right)_j \right\}\right\} \forall j \in [n]
\end{align}

Where $P^{[i]}(t) \in \mathcal{M}_{n+3\times n+3}(\R)$ and $s_f^{[i]}(t)\in \R^{n+3}$ are the solution to the differential equations:
\begin{align}
\dot{P}^{[i]} &= -C^T\left(X^{[i-1]}(t)\right)QC\left(X^{[i-1]}(t)\right) - P^{[i]}A\left (X^{[i-1]}(t)\right) -A^\top \left( X^{[i-1]}(t)\right)P^{[i]} \\&+ P^{[i]}B\left( X^{[i-1]}(t) \right)R^{-1}B^\top\left(X^{[i-1]}(t)\right)P^{[i]} \\
P^{[i]}(t_f) &= C^\top \left( X^{[i-1]}(t_f) \right) F C \left( X^{[i-1]}(t_f) \right)
\end{align}

\begin{align}
\dot{s_f^{[i]}} &= - C^\top\left(X^{[i-1]}(t)\right)Qz(t)- \left[A\left(X^{[i-1]}(t)\right) -B\left(X^{[i-1]}(t)\right)R^{-1}B^\top \left(X^{[i-1]}(t)\right)P^{[i]}(t) \right]^\top s_f^{[i]} \\
s_f^{[i]}(t_f) &= C^\top\left(X^{[i-1]}(t_f)\right)Fz(t_f)
\end{align}

Replacing the matrices of our problem

\begin{align}
\dot{P}^{[i]}(t) &= -\begin{bmatrix}
Q & 0_{n\times 3} \\ 0_{3\times n} & 0_{3\times 3}
\end{bmatrix}- P^{[i]}A\left (X^{[i-1]}(t)\right) -A^\top \left( X^{[i-1]}(t)\right)P^{[i]} + \frac{1}{\lambda}P^{[i]}B\left( X^{[i-1]}(t) \right)B^\top\left(X^{[i-1]}(t)\right)P^{[i]} \\
P^{[i]}(t_f) &= \begin{bmatrix}
0_{n\times n} & 0_{n\times 3} \\ 0_{3\times n} & 0_{3\times 3}
\end{bmatrix}
\end{align}

\begin{align}
\dot{s_f}^{[i]}(t) &= -\begin{bmatrix}Qz(t) \\ 0_{3\times 1} \end{bmatrix}- \left[A\left(X^{[i-1]}(t)\right) -\frac{1}{\lambda}B\left(X^{[i-1]}(t)\right)B^\top \left(X^{[i-1]}(t)\right)P^{[i]}(t) \right]^\top s_f^{[i]} \\
s_f^{[i]}(t_f) &=   \begin{bmatrix}
0_{n\times n}  & 0_{n\times 3} \\
\end{bmatrix}^\top z(t_f)
\end{align}

For certain entries of the dynamic the constantly zero function is a solution for them, implying by existence and uniqueness that they should be constantly zero. Then $P^{[i]}$ has $n\times n$ non zero entries and $s^{[i]}$ has $n$ non zero entries, explicitly:

\begin{align}
P^{[i]}(t) = \begin{bmatrix}
\tilde{P}^{[i]}(t) & 0_{n\times 3} \\ 0_{3 \times n} & 0_{3 \times 3}
\end{bmatrix} \\
s_f^{[i]}(t) = \begin{bmatrix}
\tilde{s}_f^{[i]}(t) \\ 0_{1\times 3}
\end{bmatrix} \\
A\left (X^{[i-1]}(t)\right) = \begin{bmatrix}
A_{11} & A_{12} \\ A_{21} & A_{22}
\end{bmatrix} \\
B\left (X^{[i-1]}(t)\right) = \begin{bmatrix}
B_1 \\ B_2
\end{bmatrix} 
\end{align}

Where $A_{11} = \diag \left( f\left(s^{[i-1]}\right)- D_{n\times 1} \right)$, and $B_1 = \diag\left(f\left(s^{[i-1]}\right)\right)\diag \left(x^{[i-1]} \right) $ The equations for $\tilde{P}^{[i]}(t)$:
\begin{align}
\dot{\tilde{P}}^{[i]}(t)= -Q- \tilde{P}^{[i]}A_{11} -A_{11}^\top \tilde{P}^{[i]} + \dfrac{1}{\lambda} \tilde{P}^{[i]}B_1B_1^\top \tilde{P}^{[i]} \\
\tilde{P}^{[i]}(t_f) = 0
\end{align}

Inspecting the former equation one notices that if $i\neq j$, $P_{ij}(t) = 0$ is a solution of all non diagonal entries. And therefore by existence and uniqueness they should be constantly zero. Hence only the diagonal entries should be calculated. 
\begin{align}
\dot{\tilde{P}}^{[i]}_{jj}(t)= -Q_{jj}- 2 \left(f_j\left(s^{[i-1]}\right)- D \right)\tilde{P}^{[i]}_{jj} + \dfrac{1}{\lambda}\left(\tilde{P}^{[i]}_{jj}\right)^2f_j\left(s^{[i-1]}\right)^2\left(x^{[i-1]}_j\right)^2  \\
\tilde{P}^{[i]}_{jj}(t_f) = 0
\end{align}

For and  $\tilde{s_f}^{[i]}(t)$ the system reduces to:

\begin{align}
\dot{\tilde{s}}_f^{[i]}(t) &= -z(t)- \left[A_{11} -\frac{1}{\lambda}B_1B_1^\top \tilde{P}^{[i]}(t) \right]^\top \tilde{s}_f^{[i]} \\
\tilde{s}_f^{[i]}(t_f) &= 0
\end{align}

And the control law is given by
\begin{align}
v^{[i]}(t)_j = \max \left\{ -1,\min\left\{0,\left( -\dfrac{1}{\lambda}B_1^\top\left(X^{[i-1]}(t)\right)\left(\tilde{P}^{[i]}(t)x^{[i]}(t)-\tilde{s}_f^{[i]}(t)\right)\right)_j \right\}\right\} \forall j \in [n]
\end{align}

They were solved using standard backward numerical integration.

\bibliographystyle{elsarticle-num}
\bibliography{mybibfile}
\end{document}